systemfit: A Package for Estimating\\ Systems of Simultaneous Equations in \proglang{R}

### 2007-07-23 paper/trunk/systemfit_intro.tex
renamed section "Testing reliability" to "Replication of textbook results"; added more goals of this section (teaching + reproducible research)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@776 591db2bd-e0a0-4d22-8c19-2192dbb7da31
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@817 591db2bd-e0a0-4d22-8c19-2192dbb7da31

% This paper is organized as follows:
In Section~\ref{sec:statistics} we introduce the statistical
background of estimating equation systems.
The implementation of the statistical procedures in \proglang{R} is
shortly explained in Section~\ref{sec:code}.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit} and how some of the features presented in the previous
section can be utilized.
In Section~\ref{sec:reliability} the reliability of the results from
\pkg{systemfit} are presented.
Finally, a summary and outlook are presented in
Section~\ref{sec:Summmary}.

>>>>

% This paper is organized as follows:
In Section~\ref{sec:statistics} we introduce the statistical
background of estimating equation systems.
The implementation of the statistical procedures in \proglang{R} is
briefly explained in Section~\ref{sec:code}.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit} and how some of the features presented in the previous
section can be used.
In Section~\ref{sec:reliability} we replicate several textbook results
with the \pkg{systemfit} package.
Finally, a summary and outlook are presented in
Section~\ref{sec:Summmary}.

### 2007-07-23 paper/trunk/systemfit_restrictions.tex
replaced all \widehat{X} by the macro \XHat
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@721 591db2bd-e0a0-4d22-8c19-2192dbb7da31
replaced all \widehat{\lambda} by the macro \lHat, which has been defined as \hat{\lambda}
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@722 591db2bd-e0a0-4d22-8c19-2192dbb7da31
increased space between rows in some arrays
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@723 591db2bd-e0a0-4d22-8c19-2192dbb7da31
renamed (subsub)section that describes the calculation of restricted estimators (method 1 and 2) -> Calculation

In many empirical applications,
it is desirable to impose restrictions on the parameters
that are derived from the underlying theoretical model.%
\footnote{%
For instance in econometric demand and production analysis,
it is common to impose (and test) homogeneity and symmetry properties.
}
There are two ways to impose linear parameter restrictions.
First, a matrix $M$ can be specified that
\begin{equation}
   \beta = M \cdot \beta^M \label{eq:T-restr} ,
\end{equation}
where $\beta^M$ is a vector of restricted (linear independent) coefficients,
and $M$ is a matrix with the number of rows equal to the number of
unrestricted coefficients ($\beta$) and
the number of columns equal to the number of restricted coefficients
($\beta^M$).
$M$ can be used to map each unrestricted coefficient to one or more
restricted coefficients.

The second way to impose linear parameter restrictions
can be formulated by
\begin{equation}
   R \beta^R = q ,
   \label{eq:restr-R}
\end{equation}
where $\beta^R$ is the vector of the restricted coefficients,
and $R$ and $q$ are a matrix and vector, respectively,
to impose the restrictions \citep[see][p.\ 100]{greene03}.
Each linear independent restriction is represented by one row of $R$
and the corresponding element of~$q$.

The first way is less flexible than this latter one%
\footnote{
While restrictions like $\beta_1 = 2 \beta_2$ can be imposed by
both methods,
restrictions like $\beta_1 + \beta_2 = 4$ can be imposed only
by the second method.
}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g.\ (near) singularity of the matrices to be inverted),
while the first does not.

@@

\subsubsection[Implementation in systemfit]{Implementation in \pkg{systemfit}}

@@

The OLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{OLS} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' y \\ q 
   \end{array} \right] ,
\end{equation}
where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{OLS} \\ \widehat{\lambda}
   \end{array} \right] 
   = \sHat^2 
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\sHat$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{OLS} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The WLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{WLS} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \OHat^{-1} y \\ q
   \end{array} \right] ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{WLS} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}

@@

The SUR estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{SUR} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \OHat^{-1} y \\ q
   \end{array} \right] ,
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{SUR} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}

@@

The 2SLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{2SLS} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' y \\ q 
   \end{array} \right] .
   \label{eq:beta2SLSr}
\end{equation}
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{2SLS} \\ \widehat{\lambda}
   \end{array} \right] 
   = \sHat^2 
   \left[ \begin{array}{cc}
      \widehat{X}' \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\sHat^2$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{2SLS} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \OHat^{-1} \widehat{X} & R' \\
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The W2SLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{W2SLS} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \OHat^{-1} \widehat{X} & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \OHat^{-1} y \\ q
   \end{array} \right] ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{W2SLS} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \OHat^{-1} \widehat{X} & R' \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}

@@

      \bHat^R_{3SLS} \\ \widehat{\lambda}

@@

      \widehat{X}' \OHat^{-1} \widehat{X} & R' \\

@@

      \widehat{X}' \OHat^{-1} y \\ q

@@

      \bHat^R_{3SLS-IV} \\ \widehat{\lambda}

@@

      \widehat{X}' \OHat^{-1} X & R' \\

@@

      \bHat^R_{3SLS-GMM} \\ \widehat{\lambda}

@@

      \bHat^R_{3SLS-Schmidt} \\ \widehat{\lambda}

@@

      \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' y \\ q

@@

      \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
      H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X} & 0' \\

@@

      \bHat^R_{3SLS-EViews} \\ \widehat{\lambda}

@@

      \widehat{X}' \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)

>>>>

In many empirical applications,
it is desirable to impose restrictions on the parameters
that are derived from the underlying theoretical model.%
\footnote{%
For instance in econometric demand and production analysis,
it is common to impose (and test) homogeneity and symmetry properties.
}
There are two methods to impose linear parameter restrictions.
First, a matrix $M$ can be specified that
\begin{equation}
   \beta = M \cdot \beta^M \label{eq:T-restr} ,
\end{equation}
where $\beta^M$ is a vector of restricted (linear independent) coefficients,
and $M$ is a matrix with the number of rows equal to the number of
unrestricted coefficients ($\beta$) and
the number of columns equal to the number of restricted coefficients
($\beta^M$).
$M$ can be used to map each unrestricted coefficient to one or more
restricted coefficients.

The second method to impose linear parameter restrictions
can be formulated by
\begin{equation}
   R \beta^R = q ,
   \label{eq:restr-R}
\end{equation}
where $\beta^R$ is the vector of the restricted coefficients,
and $R$ and $q$ are a matrix and vector, respectively,
to impose the restrictions \citep[see][p.\ 100]{greene03}.
Each linear independent restriction is represented by one row of $R$
and the corresponding element of~$q$.

The first method is less flexible than the second%
\footnote{
While restrictions like $\beta_1 = 2 \beta_2$ can be imposed by
both methods,
restrictions like $\beta_1 + \beta_2 = 4$ can be imposed only
by the second method.
}, 
and is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g.\ (near) singularity of the matrices to be inverted),
while the first does not.

@@

\subsubsection{Calculation of restricted estimators}

@@

The OLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{OLS} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' y \\ q 
   \end{array} \right] ,
\end{equation}
where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{OLS} \\[0.2em] \lHat
   \end{array} \right] 
   = \sHat^2 
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\sHat$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{OLS} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The WLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{WLS} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \OHat^{-1} y \\ q
   \end{array} \right] ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{WLS} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}

@@

The SUR estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{SUR} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \OHat^{-1} y \\ q
   \end{array} \right] ,
\end{equation}
where $\OHat = \SHat \otimes I$ and
$\SHat$ is the estimated covariance matrix of the disturbance terms.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{SUR} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \OHat^{-1} X & R' \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}

@@

The 2SLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{2SLS} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \XHat' \XHat & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \XHat' y \\ q
   \end{array} \right] .
   \label{eq:beta2SLSr}
\end{equation}
If the whole system is treated as one single equation,
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{2SLS} \\[0.2em] \lHat
   \end{array} \right] 
   = \sHat^2 
   \left[ \begin{array}{cc}
      \XHat' \XHat & R' \\
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\sHat^2$ is the estimated variance of the disturbance terms.
If the disturbance terms of the individual equations
are allowed to have different variances, 
an estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{2SLS} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \XHat' \OHat^{-1} \XHat & R' \\
      R & 0
   \end{array} \right]^{-1} ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.

@@

The W2SLS estimator restricted by $R \beta^R = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \bHat^R_{W2SLS} \\[0.2em] \lHat
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \XHat' \OHat^{-1} \XHat & R' \\
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \XHat' \OHat^{-1} y \\ q
   \end{array} \right] ,
\end{equation}
where $\OHat = \SHat \otimes I$,
$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
$\sHat_{ii} = \sHat_i^2$ is the estimated variance
of the disturbance term in the $i$th equation.
An estimator of the covariance matrix of the estimated parameters is
\begin{equation}
   \COVHat
   \left[ \begin{array}{c}
      \bHat^R_{W2SLS} \\[0.2em] \lHat
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \XHat' \OHat^{-1} \XHat & R' \\
      R & 0
   \end{array} \right]^{-1} .
\end{equation}

@@

      \bHat^R_{3SLS} \\[0.2em] \lHat

@@

      \XHat' \OHat^{-1} \XHat & R' \\

@@

      \XHat' \OHat^{-1} y \\ q

@@

      \bHat^R_{3SLS-IV} \\[0.2em] \lHat

@@

      \XHat' \OHat^{-1} X & R' \\

@@

      \bHat^R_{3SLS-GMM} \\[0.2em] \lHat

@@

      \bHat^R_{3SLS-Schmidt} \\[0.2em] \lHat

@@

      \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' y \\ q

@@

      \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
      H \left( H' H \right)^{-1} H' \OHat^{-1} \XHat & 0' \\

@@

      \bHat^R_{3SLS-EViews} \\[0.2em] \lHat

@@

      \XHat' \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)

### 2007-07-23 paper/trunk/systemfit_summary.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@817 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\subsubsection*{Unbalanced datasets}
Currently, the \pkg{systemfit} package requires
that all equations have the same number of observations.
However, many data sets have unbalanced observations.%
\footnote{
For instance,
forestry datasets typically contain many observations of inexpensive
variables (stem diameter, tree count) and few expensive variables such
as stem height or volume.
}
Simply dropping data points that do not contain observations for all
equations may reduce the number of observations considerably, and
thus, the information utilized in the estimation.
Hence, it is our intention to include the capability for estimations
with unbalanced data sets as described in \citet{schmidt77} in future
releases of \pkg{systemfit}.

>>>>

\subsubsection*{Unbalanced datasets}
Currently, the \pkg{systemfit} package requires
that all equations have the same number of observations.
However, many data sets have unbalanced observations.%
\footnote{
For instance,
forestry datasets typically contain many observations of inexpensive
variables (stem diameter, tree count) and few expensive variables such
as stem height or volume.
}
Simply dropping data points that do not contain observations for all
equations may reduce the number of observations considerably, and
thus, the information utilized in estimation of system parameters.
Hence, it is our intention to include the capability for estimations
with unbalanced data sets as described in \citet{schmidt77} in future
releases of \pkg{systemfit}.
